---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyr)
```

### Load data

```{r load-data}
load("~/Data Science Project 1/4tiY2fqCQa-YmNn6gnGvzQ_1e7320c30a6f4b27894a54e2de50a805_brfss2013.RData")
```

* * *

## Part 1: Data

The data within this file is a broad spectrum of data that is pulled from a random convenience sample of person's over the age of 18. Due to the confounding factors of the population needing to have a cell phone within a private residence or college campus, there is a risk of a portion of the youth population not being represented within this study. The measures of the data allow for a generalizable conclusion to be provided, however, due to the fact random sampling wasn't imparted causation isn't possible. 

* * *

## Part 2: Research questions

**Research quesion 1:**
Does the level of schooling completed positively affect the ability to generate income above 75k per year?

**Research quesion 2:**
Does income amount affect the average number of days of depression felt over the course of a month? 

**Research quesion 3:**
Is the average height of basketball players greater than that of baseball/ softball and soccer?

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}
Q1 <- brfss2013 %>% group_by(educa) %>% summarise( income2 > 75000) %>% count()

Q1 <- Q1[-7,]

ggplot(Q1, aes(x = educa , y = n, fill= educa )) + geom_col() + scale_x_discrete(label=function(x) abbreviate(x, minlength=1)) + theme_dark() + xlab( "Education") + ylab(" # of people above $75K") 

```



**Explanation  1:**

The data shows that the greater amount of time within the school system there is a greater ability to earn above the metric of $75k per year. The interesting thing to notice is there being a dip within the 1-3 year mark of college, given the fact that that some of the surveyed population would be in school learning their vocation in order to improve earning potential. In contrast, high school graduates could potentially have more time within their career which would allow them to make more money. 
  



**Research quesion 2:**

```{r}
Q2 <- brfss2013 %>% group_by(., income2) %>% summarise_at(., vars(qlmentl2), list(mean), na.rm = T)

Q2 <- Q2[-9,]

ggplot(Q2, aes(income2, qlmentl2, fill = income2)) + geom_col() + theme_dark() +  xlab( "Income") + ylab("  Average Number of Days Depressed Per Month ") + scale_x_discrete(label=function(x) abbreviate(x, minlength=1))
```



**Explanation  2:**
  Does money buy happiness, according the data to an extent. There is a drastic drop in the average number of day depression felt over the course of the month ones the $30K threshold is surpassed. This could potentially be attuned to the lesser need to worry about attaining good that are needed for survival and having some amount of expendable income. 


**Research quesion 3:**

```{r}
Q3 <- brfss2013 %>% summarise(., exract11, htin4) %>% group_by(exract11) %>% filter(., exract11 %in% c("Basketball", "Softball/Baseball", "Soccer"))

ggplot(Q3, aes(x = exract11, y = htin4, fill = exract11)) + geom_boxplot() + theme_dark() + xlab( "Sport ") + ylab(" Height ") 



```



**Explanation  3:**
  Is there a difference in the average height of athletes pertaining  to the sports of basketball, softball/baseball, and soccer? Yes, there is. Given the individual needs for longer limbs or taller heights to gain an advantage over one's opponent taller people would be favored in the sport of basketball. Basketball and softball favor long limb in order to create power throwing the ball. Soccer favors maneverabily therefore smaller, more agile indivduals may have the advantage   

